---
title: "Tutorial of PRECISION-LVAD!"
output: learnr::tutorial
# progressive: true
# allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(qcc)
library(dplyr)
library(lme4)
library(utils)
library(readxl)
tutorial_options(
  exercise.timelimit = 60,
  # A simple checker function that just returns the message in the check chunk
  exercise.checker = function(check_code, ...) {
    list(
      message = eval(parse(text = check_code)),
      correct = logical(0),
      type = "info",
      location = "append"
    )
  }
)
knitr::opts_chunk$set(error = TRUE)


patient_specific_residual<-  function(variable_name, df_all_patient){
  
  
  # This function takes a variable name and a data frame containing all patient data 
  # and returns a data frame containing residuals for each patient. 
  
  # Arguments:
  # variable_name: the name of the variable for which residuals are to be calculated.
  # df_all_patient: the data frame containing all patient data.
  
  
  
  # Find all unique patient IDs in the data frame.
  u= sort(unique(df_all_patient$ID))
  
  # Create an empty data frame to store residuals for all patients.
  df_contain_resid_all_patient<-data.frame()
  
  # Loop through all patients.
  for(k in 1:length(u)){
    
    # Subset the data frame to contain only data for the current patient.
    df_patient<- df_all_patient[df_all_patient$ID==u[k],]
    
    # Subset the admission file to contain only admission data for the current patient.
    patient_admission<- admission_file[admission_file$ID==u[k] ,]
    
    # Find all dates on which the stored speed value changed for the current patient.
    dates_speed_changed<- sort(df[which(df_patient$Stored_speed != lag(df_patient$Stored_speed)),]$Datetime)
    
    # If the stored speed value never changed, use only admission dates as events.
    if (length(dates_speed_changed)!=0) {
      events<- sort(unique(patient_admission$admission_date), 
                    unique(dates_speed_changed))
    }else{
      events<- unique(patient_admission$admission_date)
    }
    
    # Create an empty data frame to store residuals for the current patient.
    residuals<- data.frame()
    
    # Get the total number of observations for the current patient.
    number_of_obs<-nrow(df_patient)
    
    # If there are no events for the current patient, skip to the next patient.
    if (length(events)==0) {
      next
    }else{
      
      # If there are multiple events for the current patient, calculate residuals 
      # for each event.
      if (length(events)>1) {
        
        # Find the index of the last observation before the first event.
        calibration_index<- max(which(df_patient$Datetime < events[1]))
        
        # Subset the data frame to contain only observations before the first event.
        calibration_obs<- df_patient[1:calibration_index,]
        
        # Calculate residuals for the first event.
        residuals<- mean_estimator_ind_patient(df = calibration_obs, 
                                               df_normal = pws_patient, 
                                               variable_name = variable_name)
        
        # Create an empty data frame to store residuals for the current event.
        residuals_temporal<- residuals_temporal[(1:(nrow(calibration_obs))),]
        
        # Loop through all events except for the first one.
        for (t in 2:length(events)) {
          
          # Find the indices of the observations between the current and previous event.
          i<- max(which(df_patient$Datetime <= events[t]))
          j<- max(which(df_patient$Datetime <= events[t-1]))
          
          # Subset the data frame to contain only observations between the current and previous event.
          calibration_obs<- df_patient[(j+1):i,]
          
          # Calculate residuals for the current event.
          residuals_temporal<- mean_estimator_ind_patient(df = calibration_obs, 
                                                          df_normal = pws_patient, 
                                                          variable_name = variable_name)
          
          # Create an empty data frame to store residuals for the current event.
          residuals_temporal<- residuals_temporal[(1:(nrow(calibration_obs))),]
          
          # Add the residuals for the current event to the residuals
          
          residuals<- rbind(residuals, residuals_temporal)
        }  
        
        
        
        number_of_resid<- nrow(residuals)
        
        
        if (number_of_resid < number_of_obs) {
          
          calibration_obs<- df_patient[(number_of_res+1):number_of_obs,]
          
          residuals_temporal<-  mean_estimator_ind_patient(df = calibration_obs, 
                                                           df_normal = pws_patient, 
                                                           variable_name = variable_name)
          
          residuals_temporal<- residuals_temporal[(1:(nrow(calibration_obs))),]
          residuals<- rbind(residuals, residuals_temporal)
          
          
        }
        
        
        
      }else{
        calibration_index<- max(which(df_patient$Datetime < events[1]))
        
        calibration_obs<- df_patient[1:calibration_index,]
        
        residuals_temporal<- mean_estimator_ind_patient(df = calibration_obs, 
                                                        df_normal = pws_patient, 
                                                        variable_name = variable_name)
        
        
        
        residuals<- rbind(residuals, residuals_temporal)
        
        
        number_of_res<- nrow(residuals)
        
        if (number_of_res < number_of_obs) {
          
          
          calibration_obs<- df_patient[(number_of_res+1):number_of_obs,]
          
          
          residuals_temporal<- mean_estimator_ind_patient(df = calibration_obs, 
                                                          df_normal = pws_patient, 
                                                          variable_name = variable_name)
          
          
          residuals_temporal<- residuals_temporal[(1:(nrow(calibration_obs))),]
          
          residuals<- rbind(residuals, residuals_temporal)
          
          
          
          
        }
        
        
        
      }
    }
    
    
    
    
    
    df_patient<- cbind(df_patient, residuals)
    
    
    
    
    df_contain_resid_all_patient<-rbind(df_contain_resid_all_patient, df_patient)
    
  }
  
  
  return(df_contain_resid_all_patient)
  
  
}
mean_estimator_ind_patient<- function(df,
                                      df_normal,
                                      variable_name){
  
  df$calibrated_resid<-0
  iter<-180
  kl<-nrow(df)
  
  if (iter> kl) {
    iter<-kl
  }
  
  bank_data<- df_normal %>%group_by(ID)%>%slice(1:28)
  bank_data$Datetime<- as.Date(bank_data$Datetime)
  bank_data$calibrated_resid<-0
  bank_data$mean_estimation<-0
  df$mean_estimation<-0
  online_mean<-data.frame()
  
  patient_num = df$ID[1]
  if (variable_name == "Flow") {
    
    
    for (j in 1:(iter)) {
      
      a<- df[1:j,]
      
      new_df<-data.frame()
      new_df<- rbind(bank_data, a)
      
      
      
      
      
      
      new_model<- lmer(Flow ~  (1 + norm_speed |ID) + HCT 
                       , data = new_df, REML = F, 
                       control = lmerControl(optimizer ="bobyqa", calc.derivs = F))
      
      
      the_estimated_mean<- 
        coef(new_model)$ID[which( row.names(coef(new_model)$ID)==patient_num),3]* a$HCT+
        a$norm_speed *coef(new_model)$ID[which( row.names(coef(new_model)$ID)==patient_num),1]+
        coef(new_model)$ID[which( row.names(coef(new_model)$ID)==patient_num),2]
      
      
      online_mean<- rbind(online_mean, the_estimated_mean)
      
      df[j,"calibrated_resid"]<- as.numeric(df$Flow[j]- online_mean[j,1])
      df[j,"mean_estimation"]<-  online_mean[j,1]
      
      
    }
    
    if (kl>iter) {
      df[((iter+1): kl), "calibrated_resid"]<- df[((iter+1):kl), "Flow"] - online_mean[iter,1]
      df[((iter+1): kl), "mean_estimation"]<- online_mean[iter,1]
    }
    
    
    df$sigma<-0
    
    for (s in 1:iter) {
      
      df$sigma[s]<- sqrt(var(df[1:s,"calibrated_resid"])[1])
      
      
    }
    df$sigma[1]<-2
    
    if (kl>iter) {
      df[((iter+1): kl), "sigma"]<- df[iter,"sigma"]
    }
    
  }else if(variable_name == "Motor_power"){
    for (j in 1:(iter)) {
      
      a<- df[1:j,]
      
      new_df<- rbind(bank_data, a)
      
      new_model<- lmer(Motor_power ~  (1 + norm_speed |ID) 
                       
                       , data = new_df, REML = F, 
                       control = lmerControl(optimizer ="bobyqa", calc.derivs = F))
      
      
      the_estimated_mean<- coef(new_model)$ID[which( row.names(coef(new_model)$ID)==patient_num),2]+
        a$norm_speed *coef(new_model)$ID[which( row.names(coef(new_model)$ID)==patient_num),1]
      
      
      online_mean<- rbind(online_mean, the_estimated_mean)
      
      df[j,"calibrated_resid"]<- as.numeric(df$Motor_power[j]- online_mean[j,1])
      df[j,"mean_estimation"]<-  online_mean[j,1]
      
      
    }
    df[((iter+1): kl), "calibrated_resid"]<- df[((iter+1):kl), "Motor_power"] - online_mean[iter,1]
    df[((iter+1): kl), "mean_estimation"]<- online_mean[iter,1]
    
    df$sigma<-0
    
    for (s in 1:iter) {
      
      df$sigma[s]<- sqrt(var(df[1:s,"calibrated_resid"])[1])
      
      
    }
    df$sigma[1]<-2
    df[((iter+1): kl), "sigma"]<- df[iter,"sigma"]
    
    
  }
  
  
  
  df$updated<-0
  df$updated[1]<-1
  control_limits<- data.frame(lcl = rep(0, iter), ucl = rep(0,iter), mean_estimate = rep(0, iter))
  
  
  
  return(df[, c("calibrated_resid", "mean_estimation","sigma", "updated")])
}


```

## Welcome

Welcome to PRECISION-LVAD tutorial!In this tutorial, we'll be learning :

* Load datasets
* Fit personalized model on each pump parameter based on `PRECISION-LVAD` engine
* Get `Out-of-control` points for each time-point
* Evaluate the results

This srepository is linked to our new paper entitled:[paper](http://www.exmaple.com)

### Prerequisites
To run the PRECISION-LVAD algorithm you need to install below packages:

* `dplyr`
* `Lme4`
* `qcc`




## Load data
```{r, filter1, exercise = TRUE, exercise.setup= "one"}


# loading the datasets:
df_all_patient<-readr::read_csv("data/syn_data_CA.csv")
head(df_all_patient)
```

```{r, filter2, exercise = TRUE, exercise.setup= "one"}


# loading the datasets:
df_all_patient<-readr::read_csv("data/syn_data_MB.csv")
head(df_all_patient)
```


```{r, filter3, exercise = TRUE, exercise.setup= "one"}


# loading the datasets:
admission_file<-readr::read_csv("data/admission_file.csv")
head(admission_file)
```

```{r, filter4, exercise = TRUE, exercise.setup= "one"}


# loading the datasets:
pws_patient<-readr::read_csv("data/stable_patient.csv")
head(pws_patient)
```

```{r, filter5, exercise = TRUE, exercise.setup= "one"}
df_all_patient$Datetime<- as.Date(df_all_patient$Datetime)
pws_patient$Datetime<- as.Date(pws_patient$Datetime)
```


## Fit PRECISION-LVAD on the data

```{r, filter6, exercise = TRUE, exercise.setup= "one"}

df<-patient_specific_residual(variable_name = "Flow",
                              df_all_patient = df_all_patient)

head(df)
```
## Obtain Out-of-control points with PRECISION-LVAD

## Evaluate

